---
title: r u bot or not bot?
author: William Bourgeois
date: '2019-09-27'
slug: r-u-bot-or-not-bot
categories: []
tags:
  - Twitter
  - Network analysis
---

If you have a Twitter account chances are one day you had a new follower that you did not know and that felt a bit odd. Maybe the account had few followers, a lot or no tweets, had been recently created, is not followed by anyone you're following, or likely a combination of the above.  

It is also likely that these accounts are managed by bots using the Twitter API. Since it is estimated that up to 15% of profiles on Twitter are controlled by bots and that some of them are said to have political impact through their propagation of propaganda, there has been a large number of studies and analysis into bots and more specifically bots re-tweeting trolls. 

A whole range of tools can be applied to study bots: supervised learning, network & cluster analysis, time series, NLP...  

But in this post I will try to detect and analyse bots following a randomly chosen Twitter friend. 

Please note that in Twitter parlance a follower is someone who is following your account and a friend is an account you follow. 

Let's get some data on who follows @kderderian1 (kd) using the rtweet package!

```{r message=FALSE, warning=TRUE}
library("tidyverse")
library("rtweet")
library("glue")
library("visNetwork")
```


```{r}
followers_id <- get_followers("kderderian1", n = 5000)  # all the followers
followers_data <- lookup_users(followers_id$user_id)    # data on the accounts
followers_data %>% select(screen_name, text) %>%        # a sample
  mutate(text = substr(text, 1, 60)) %>% 
  filter(!is.na(text)) %>% 
  slice(10:20)
```

There is a package ("tweetbotornot") that is designed to analyse Twitter accounts and that returns a probability score for being a bot. (I had some problems running it but managed to debug a copy I made on a github repo and installed it from there.)

Let's run it on the followers. Since we have to analyse a few hundreds of them, we need to take into account the Twitter API rate limits. So we can use a loop to do just that.

```{r eval=FALSE}
library("glue")
library("tweetbotornot")

data <- data.frame()
not_auth <- data.frame()
limits <- rate_limits()

for (i in followers_data$screen_name){
  print(i)
  
  row_n <- match(i, followers_data$screen_name)
  print(row_n)
  
  row <- try(botornot(i))
  
  if(inherits(row,"try-error")) {
    
    not_auth <- rbind(not_auth, as.data.frame(i))
    next}
  
  if ((row_n %% 10) == 0) {
     limits <- rate_limits()
     print(glue("{limits$query[49]} = {limits$remaining[49]}"))
     print(glue("{limits$query[11]} = {limits$remaining[11]}"))
   }
   
  if (limits$remaining[11] < 20) {
     Sys.sleep(as.numeric((limits$reset_at[11] - limits$timestamp[11])*60)+10)
     limits <- rate_limits()
   }
   
  if (limits$remaining[49] < 20) {
     Sys.sleep(as.numeric((limits$reset_at[49] - limits$timestamp[49])*60)+10)
     limits <- rate_limits()
   }
  
  data <- rbind(data, row)
}

 
```


```{r eval=FALSE, include=FALSE}
#Save it for posterity
#saveRDS(data, "data/20190927/data_followers_kd.rds")
#saveRDS(not_auth, "data/20190927/no_data_followers_kd.rds")
#saveRDS(followers_data, "data/20190927/followers_kd.rds")
```

```{r eval=TRUE, include=FALSE}
data_followers_kd <- readRDS("~/R/blogs/content/post/data/20190927/data_followers_kd.rds")
```

We have a bot probability calculation for 1.322 followers and were refused access to 117 follower's accounts. Some of these might of course also be bots. Like for instance 'summer_intern94', account created July 2017, no tweets, following 871, 24 followers. 

```{r}
nrow(data_followers_kd[data_followers_kd $prob_bot > 0.5,])
```

The issue is that a majority of the followers have a probability score of more than 50%. And there are a lot of false positives in the top scores. So even though it might be difficult to easily identify Twitter bots, I'm not really satisfied with the results of this package. 

There also do not seem to be a lot of alternatives. One of them though is a joint project of the Network Science Institute (IUNI) and the Center for Complex Networks and Systems Research (CNetS) at Indiana University. It uses supervised learning to compute a bot score based on about 1200 features also extracted from the Twitter API. More [here](https://botometer.iuni.iu.edu/#!/faq#how-does-it-works).

There are a number of ways to obtain results through their [website](https://botometer.iuni.iu.edu/#!/), you can for instance check all your Twitter followers, but I wanted to use their API, using a [RapidAPI](https://rapidapi.com/) account and an R client [library](https://github.com/marsha5813/Botcheck)

```{r}
#install.packages("devtools")
#library(devtools)
#install_github("marsha5813/botcheck")
library("botcheck")
# Load dependencies
library("httr")
library("xml2") 
library("RJSONIO")
library("rtweet")
library("tidyverse")
library("visNetwork")
```

You must first [get your own Botometer API key](https://market.mashape.com/OSoMe/botometer). 

You'll also need to create and register a [Twitter app](https://apps.twitter.com/). Save your keys by replacing the "xxxx" placeholders.


```{r, eval=FALSE}

Mashape_key = "xxxx"

consumer_key = "xxxxx"
consumer_secret = "xxxxx"
access_token = "xxxxx"
access_secret = "xxxxx"
```


```{r echo=TRUE}
source("~/R/blogs/content/post/data/20190927/secrets.r")

```

Finally, store the Twitter authorization info in a local file.

```{r}

Mashape_key = mashape_key

myapp = oauth_app("twitter", key= consumer_key, secret= consumer_secret)
sig = sign_oauth1.0(myapp, token= access_token, token_secret= access_secret)
```

Call the botcheck() function to get the probability that a Twitter handle is a bot. Do not put the "@" before the Twitter handles. For example:

```{r}
botcheck("barackobama")
```


```{r eval=FALSE, include=TRUE}
library("glue")
data <- data.frame()
not_auth <- data.frame()
n <- nrow(data)
m <- nrow(followers_data)
limits <- rate_limits()

for (i in followers_data$screen_name[n:m]){
  print(i)
  row_n <- match(i, followers_data$screen_name)
  print(row_n)
  
  row <- try(botcheck(i))
  
  if(length(row) ==0) next()
  
  
  row <- as.data.frame(cbind(row, i))
  
  if(inherits(row,"try-error")) {
    
    not_auth <- rbind(not_auth, as.data.frame(i))
    next}
  
  
  if ((row_n %% 10) == 0) {
     limits <- rate_limits()
     print(glue("{limits$query[49]} = {limits$remaining[49]}"))
     print(glue("{limits$query[11]} = {limits$remaining[11]}"))
   }
   
  if (limits$remaining[11] < 20) {
     Sys.sleep(as.numeric((limits$reset_at[11] - limits$timestamp[11])*60)+10)
     limits <- rate_limits()
   }
   
  if (limits$remaining[49] < 20) {
     Sys.sleep(as.numeric((limits$reset_at[49] - limits$timestamp[49])*60)+10)
     limits <- rate_limits()
   }
  
  data <- rbind(data, row)
  
  
  if ((row_n %% 30) == 0){
    saveRDS(data, "data/20190927/data.rds")
  }

}


```

```{r include=FALSE}
data_1 <- readRDS("~/R/blogs/content/post/data/20190927/data_1.rds")
data_2 <- readRDS("~/R/blogs/content/post/data/20190927/data_2.rds")
data_3 <- readRDS("~/R/blogs/content/post/data/20190927/data_3.rds")
data_4 <- readRDS("~/R/blogs/content/post/data/20190927/data_4.rds")

data <- rbind(data_1,
               data_2,
               data_3,
               data_4) %>% 
  unique() %>% 
  mutate(score = as.numeric(as.character(row))) %>% 
  arrange(score) %>% 
  mutate(range = as.numeric(rownames(.))) %>% 
  filter(!is.na(row)) %>% 
  arrange(range)
  

rm(data_1, 
   data_2, 
   data_3, 
   data_4) 

```


Again we do not have data for all the accounts, but the results look more promising. In the plot the  
"tweetbotornot" results are in blue and the "botometer" in red. The red line indeed looks more realistic. 

```{r message=FALSE}
y <- data_followers_kd %>% 
  arrange(prob_bot) %>% 
  unique() %>% 
  mutate(range = as.numeric(rownames(.)))

data_bots <- data %>% 
  left_join(y) %>% 
  mutate(row = as.numeric(as.character(row)))

ggplot(data = data_bots, aes(x = range, y= row, group = 1)) +
  geom_smooth(method = 'gam', formula  = y ~ s(x, bs = "cs"), color = "darkred") + 
  geom_line(aes(y = prob_bot), color="steelblue")  + 
  scale_x_continuous(expand = c(0, 0)) + 
  theme_minimal()

```

Now let's analyse part of the network and also have look at the friends of the followers (or put differently who else do the followers follow).

```{r message=FALSE, results='hide', eval=FALSE}
followers <- readRDS("~/R/blogs/content/post/data/20190927/data_followers_kd.rds")
data <- data.frame()
limits <- rate_limits()
for(i in followers$user_id){
  print(i)
  row_n <- match(i, followers_id$user_id)
  print(row_n)
  
  if ((row_n %% 15) == 0) {
     saveRDS(data, "data_friends_1.rds")
     limits <- rate_limits()
     print(glue("{limits$query[11]} = {limits$remaining[11]}"))
     print(glue("{limits$query[98]} = {limits$remaining[98]}"))
   }
  
  
  friends_i <- get_friends(i, n = 5000, page = "-1", 
                             retryonratelimit = TRUE, 
                             parse = TRUE, verbose = TRUE, 
                             token = NULL)
  
  
  if (limits$remaining[11] < 20) {
     Sys.sleep(as.numeric((limits$reset_at[11] - limits$timestamp[11])*60)+10)
     limits <- rate_limits()
   }
  
  if (limits$remaining[98] < 2) {
     Sys.sleep(as.numeric((limits$reset_at[98] - limits$timestamp[98])*60)+10)
     limits <- rate_limits()
   }
  
  if (length(friends_i) == 0) next()
  
  row <- cbind(i,friends_i)
  data <- rbind(data, row)
  
}

#saveRDS(data, "~/R/blogs/content/post/data/20190927/data_followers_friends.rds")

```

Get the friends of kd.

```{r}
friends <- get_friends("kderderian1", n = 5000)
```

And who the friends follow (the friends of friends).

```{r message=FALSE, results='hide', eval=FALSE}
data <- data.frame()
limits <- rate_limits()
for(i in friends$user_id[826:2147]){
  print(i)
  row_n <- match(i, friends$user_id)
  print(row_n)
     limits <- rate_limits()
     print(glue("{limits$query[11]} = {limits$remaining[11]}"))
     print(glue("{limits$query[96]} = {limits$remaining[96]}"))
  
  
  friends_i <- get_friends(i, n = 5000, page = "-1", 
                             retryonratelimit = TRUE, 
                             parse = TRUE, verbose = TRUE, 
                             token = NULL)
  
  if (limits$remaining[11] < 20) {
     Sys.sleep(as.numeric((limits$reset_at[11] - limits$timestamp[11])*60)+10)
     limits <- rate_limits()
   }
  
  if (limits$remaining[96] < 2) {
     Sys.sleep(as.numeric((limits$reset_at[98] - limits$timestamp[98])*60)+10)
     limits <- rate_limits()
   }
  
  if (length(friends_i) == 0) next()
  
  row <- cbind(i,friends_i)
  data <- rbind(data, row)
  
}

#saveRDS(data, "~/R/blogs/content/post/data/20190927/data_friends_of_friends.rds")

```


```{r include=FALSE}
# data_friends_1 <- readRDS("~/R/blogs/content/post/data/20190927/data_friends_1.rds")
# data_friends_2 <- readRDS("~/R/blogs/content/post/data/20190927/data_friends_2.rds")
# data_friends_3 <- readRDS("~/R/blogs/content/post/data/20190927/data_friends_3.rds")
#  # 
# data_followers_friends <- rbind(data_friends_1,
#                         data_friends_2,
#                         data_friends_3)
#   
# rm(data_friends_1,
#      data_friends_2,
#      data_friends_3)
# 
# saveRDS(data_followers_friends, "~/R/blogs/content/post/data/20190927/data_followers_friends.rds")

data_friends_of_friends <- readRDS("~/R/blogs/content/post/data/20190927/data_friends_of_friends.rds")
data_followers_friends <- readRDS("~/R/blogs/content/post/data/20190927/data_followers_friends.rds")

```

There are almost 3 million links between friends and friend of friends. And a bit more than two million links between followers and their friends. The network needs to be limited to be plotted. 

I'll only include friends of kd in the network of friends of friends and limit the number of accounts and only take accounts with the highest probabilities of being a bot.

```{r}
data <- data_friends_of_friends %>% 
  filter(user_id %in% friends$user_id) #only include direct friends

data %>% group_by(user_id) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

```

We get about 200k links between the 2k nodes. We can get their screen names.

```{r}

users <- lookup_users(unique(data$user_id))

```


```{r message=TRUE, warning=FALSE}
# top_n <- 2206 
top_n <- 400 

top_friends <- data_friends_of_friends %>% 
  filter(user %in% friends$user_id) %>% group_by(user_id) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  slice(1:top_n)


edges_kd <- friends %>% 
  mutate(from = 3070002117) %>% 
  rename("to" = "user_id") %>% 
  select(from, to)

#edges followers of kd for selecting the bots
top_bots_user_id  <- data_bots %>% 
  filter(score > 0.975) %>% 
  select(i, score) %>% 
  rename("screen_name" ="i") %>% 
  mutate(screen_name = as.character(screen_name)) %>% 
  unique() %>% 
  left_join(data_followers_kd) %>% 
  select(-prob_bot)

bots <- data_followers_friends %>% 
  filter(user %in% top_bots_user_id$user)

edges_bots <- bots %>% 
  rename("from" = "user",
         "to" = "user_id") %>% 
  select(from, to) %>% 
  filter(to %in% c(top_friends$user_id, 3070002117))

edges <- data_friends_of_friends %>% 
  filter(user %in% top_friends$user_id) %>% 
  filter(user_id %in% top_friends$user_id) %>% 
  select(-i) %>% 
  rename("from" = "user", "to" = "user_id") %>% 
  rbind(edges_kd, edges_bots)

nodes <- c(edges$from,edges$to) %>% 
  as.data.frame() %>% 
  unique() %>% 
  rename("id" =".")

screen_names <- users %>% 
  select(user_id, screen_name)

# and get the bot screen_names fresh

bots_screen_names <- lookup_users(top_bots_user_id$user_id) %>%
  select(user_id, screen_name)

kd_screen_name  <- lookup_users(3070002117) %>%
  select(user_id, screen_name)

screen_names <- screen_names %>% 
  rbind(bots_screen_names, kd_screen_name)

nodes <- nodes %>% 
  left_join(screen_names, by = c("id" = "user_id"))

row.names(nodes) <- c()
nodes <- nodes %>% 
  rownames_to_column("index") %>% 
  mutate(index = as.character(index)) %>% 
  mutate(id = as.character(id))

edges <- edges %>% 
  mutate(from = as.character(from)) %>% 
  rename("id" = "from") %>% 
  left_join(nodes) %>% 
  rename("from" = "index") %>%
  select(-id, -screen_name) %>% 
  rename("id" = "to") %>% 
  left_join(nodes) %>%
  rename("to" = "index") %>% 
  select(-id, -screen_name)


nodes <- nodes %>% 
  select(-id) %>% 
  rename("id" = "index") %>% 
  rename("label" = "screen_name") %>% 
  mutate(color = case_when(label == "kderderian1" ~ "blue",
                           label %in% bots_screen_names$screen_name ~"red", 
                           TRUE ~ "lightblue"))%>% 
  mutate(value = case_when(label == "kderderian1" ~ 10,
                           label %in% bots_screen_names$screen_name ~ 10, 
                           TRUE ~ 1)) %>% 
  mutate(group = case_when(label == "kderderian1" ~ "kd edges",
                           label %in% bots_screen_names$screen_name ~ "bots edges", 
                           TRUE ~ "friends edges"))
```

The number of edges, nodes and bot accounts can be further tweeked. Here is a network visualisation with the current parameters.

```{r}

visNetwork(nodes, edges, height = "800px", width = "100%") %>%
  visOptions(selectedBy = "group", 
             highlightNearest = TRUE, 
             nodesIdSelection = FALSE) %>%
  visPhysics(stabilization = FALSE)
```


The visualisation function also takes into account the groups that we created and tends to push them slightly appart to partially avoid the network fur ball effect. Zooming in is a possibilty as is clicking and draging indiviual accounts. 

