---
title: tests with paws on aws
author: William Bourgeois
date: '2019-05-21'
slug: tests-with-paws-on-aws
categories: []
tags:
  - AWS
  - R
  - Sagemaker
  - Twitter
  - Amazon Translate
  - AWS Rekognition
  - Amazon Comprehend
---

When trying to find an interface between AWS and Rekognition in a previous post, I was unsuccessful in finding a suitable R library. But today I (partially) caught up with my email backlog and was excited to find in the newsletter from [rweekly](https://rweekly.org/), that a number of interesting R packages just came out at the beginning of the month.

As far as I understand at the moment, there are 11 packages bundled in a [kit](https://cran.r-project.org/web/packages/paws/index.html) published three days ago...

So my intent in this post is to take these for a ride, using our parliamentarian's tweets, and see what I can get out of different AWS services through R.


```{r message=FALSE, warning=FALSE}

library("tidyverse")
library("paws.machine.learning") # install.packages("paws.machine.learning")

```

Authentication information for AWS:

```{r eval=FALSE, include=FALSE}

# s3 bucket

user_key <- ""
api_key <- ""
region <- "eu-west-1"

Sys.setenv("AWS_ACCESS_KEY_ID" = AWSAccessKeyId,
           "AWS_SECRET_ACCESS_KEY" = AWSSecretKey,
           "AWS_DEFAULT_REGION" = region,
           "AWS_REGION" = region,
           "AWS_SESSION_TOKEN" = "")
```

Detecting languages with Amazon Comprehend:

```{r}

svc <- comprehend()
svc$batch_detect_dominant_language(
  "ik voelde me zo angstg als een jong dier onder vuur genomen door jagers")

```

Nice! Indeed Dutch. This returns 0.9996 confidence score.

```{r, eval= FALSE}
svc$batch_detect_sentiment(
  "ik voelde me zo angstg als een jong dier onder vuur genomen door jagers")
```

Returns an error. It needs to know which language it would be in.

```{r, eval= FALSE}
svc$batch_detect_sentiment(
  "ik voelde me zo angstg als een jong dier onder vuur genomen door jagers",
  LanguageCode = "nl")
```

Aw shoot does not do Dutch yet.

But can we translate Dutch to French?

```{r}
svc <- translate()
svc$translate_text("ik voelde me zo angstg als een jong dier onder vuur genomen door jagers", 
                   SourceLanguageCode = "nl", 
                   TargetLanguageCode ="fr")


```

Not perfect, but pretty close.

```{r}
text <- svc$translate_text("ik voelde me zo angstg als een jong dier onder vuur genomen door jagers",
                           SourceLanguageCode = "nl", 
                           TargetLanguageCode ="fr") 

sentiment_text <- text$TranslatedText

svc <- comprehend()
svc$batch_detect_sentiment(sentiment_text,
  LanguageCode = "fr")

```

Hmm. Not really neutral this text. 

Let's get some tweets. But just the texts for now.  

```{r}
tweets <- readRDS("~/R/blogs/content/post/data/20190503/tweets.rds") %>% 
  select(text)

nrow(tweets)

```

So we have 264K tweets. The text is still not cleaned. Let's see what we can get out of a small sample of them.

```{r}
set.seed(42)
sample <- sample_n(tweets, 20)
sample
```

Language detection on the sample.

```{r}

svc <- comprehend()
languages <- svc$batch_detect_dominant_language(sample$text)
list <- languages$ResultList

lang <- vector()
for (i in (1:length(list))){
  print(list[[i]][["Languages"]][[1]][["LanguageCode"]])
  lang_tweet <- list[[i]][["Languages"]][[1]][["LanguageCode"]]
  lang <- append(lang, lang_tweet)
}

```

So let's look at the results compared to the tweets.

```{r}
x <- as.data.frame(lang) %>% 
  cbind(sample) %>% 
  mutate(text = substr(text, 0, 50))

x

```

So that is more than good. Happy with this proof of concept. 

Before calling it a day I wanted to test the detect_entities function.

```{r}
svc$detect_entities(tweets$text[200011], LanguageCode = "fr")
tweets$text[200011]

```


It detected the handle, can be useful, but there are other ways to extract the handles from a tweet. 

What about this tweet:

```{r eval=FALSE, include=FALSE}
svc$detect_entities(tweets$text[264000], LanguageCode = "nl")
```


```{r eval=FALSE, include=FALSE}
tweets$text[264000]

```

Damn does not seem to work in Dutch.

So what about this one:

```{r}
svc$detect_entities(tweets$text[264000], LanguageCode = "fr")
tweets$text[264000]
```

So saying that it is in French resulted in a correct output (that might be an exceptional though, let's see)

```{r}
svc$detect_entities(tweets$text[260725], LanguageCode = "fr")

tweets$text[260725]
```


Clearly identified "Zakia Khattabi" in the tweet. Nice. Looking forward to test some more. 

But for now will see an episode of "Collateral" on Netflix and hope not to get to bed too late.

Thank you for your attention. 

Sleep tight. 

