---
title: google location tracking
author: William Bourgeois
date: '2019-03-30'
slug: google-location-tracking
categories:
  - geocoding
tags:
  - R Markdown
  - Google
  - Geocoding
---

Biking to work this week I was wondering if I had not gotten slower and if my commute was not taking longer than before. Not being a regular user of Strava or a similar app, I wondered if I could find an answer to that question using my Google location history.

So I downloaded the data, that comes in json format, and had a go at it. You can [download](https://takeout.google.com/settings/takeout) your Google location data from your Google account. 

Let's load the data and attach the libraries that are needed.

```{r message=FALSE, warning=FALSE}
library("jsonlite")
library("tidyverse")
library("lubridate")
library("ggplot2")
library("ggmap")      # devtools::install_github("dkahle/ggmap")
library("glue")
library("emo")        # devtools::install_github("hadley/emo")
library("viridis")    # install.packages("viridis")

```

```{r}
data <- fromJSON("./data/20190330/Location History.json") # extracts a list
locations <- data$locations # and the list contains a dataframe
rm(data)  # no need for this anymore



```



This [blogpost](https://shiring.github.io/maps/2016/12/30/Standortverlauf_post) from Shirin Glander was super useful and I am shamelessly stealing some of her great ideas. The time stamp needs to be converted to be readable. 
The field heading seems to be in degrees, is velocity in mph? And I don't know where the altitude reading comes from (maybe not a pressure sensor but gps triangulation?, ground elevation?).

```{r}


df <- locations %>% 
  mutate(datetime = as.POSIXct(as.numeric(timestampMs)/1000, origin = "1970-01-01")) 

sort(unique(df$heading)) #naught to 360


```

Is the velocity in miles per hour ?

```{r}
unique(df$velocity)
```


Does that mean that I ever broke the speed limit? We're not supposed to drive faster than 120 km/h in Belgium :police_car:!

```{r}

fine <- paste(round(max(unique(df$velocity), na.rm = TRUE)*1.609344,0),"km!")

glue("{fine} :scream: :scream: :scream:")


```

Could this information get me fined? Or will this type information be used in the future to determine guilt?
Are the police allowed to access this type of Google data to for instance find perps of hit & runs?

The data frame also contains lists with times stamps and estimations of activity at that time stamp.

```{r}

str(df$activity[80])

```

So if I read this correctly this object is a list of one, containing a dataframe with another two objects, a timestamp and another list of one, that contains a dataframe with 5 observations and two variables: the type of activity and the probability calculated for it. So on timestamp "1395248054966" (= 2014-03-19 17:54:14 CET) I had a 49% probability to be on foot. 

And then there is also a measure of altitude. Is it in feet or meters, and where do the negative numbers come from? Maybe these data points should be analysed a bit later.

```{r}
glimpse(unique(df$altitude)) # ? what sensor does it use?

```

```{r}
summary(unique(df$altitude))
```


Let's plot some data on maps. 

First some additional date prep.
```{r}
df <- df %>% 
  mutate(lat   = latitudeE7/1e7,
         lon   = longitudeE7/1e7,
         time  = strftime(datetime, format = "%H:%M:%S"),
         date  = date(datetime),
         year  = year(datetime),
         month = month(datetime),
         wday  = wday(datetime))
```


Working with ggmap is pretty straightforward, but you need to obtain an api key from google and enable billing. More info [here](https://developers.google.com/maps/documentation/embed/get-api-key). For obvious reasons you need to keep your key secret and safe, but additionally you can restrict the calls to your [IP](https://console.developers.google.com/apis/credentials?project=grounded-block-178714&folder&organizationId). Will have a go at OSM in a future post, promised. 

```{r message=FALSE}
secrets <- readRDS("~/R/geo/secret/secrets.rds")
key <- secrets$key[1]
register_google(key = key)

belgium <- get_map(location = 'Belgium', zoom = 8, maptype = "terrain-lines")


```

The map coverage is called the bounding box of the map and it is an attribute of the ggmap object. 

```{r}
bb <- attr(belgium, "bb")
bb
```

```{r}
x <- df %>% 
  filter(lat < 51.607 & lat > 49.37082,
         lon > 2.71487 & lon < 6.20495)

```


```{r fig.height=10, fig.width=10}

ggmap(belgium, maptype = "terrain-lines") + 
  geom_point(data = x, aes(x = lon, y = lat),
             alpha = 0.1, color = "#FC4E07", size = .65) + 
  theme(legend.position = "right") + 
  labs(
    x = "Longitude", 
    y = "Latitude", 
    title = "Location history data points in Belgium",
    caption = "\nA simple point plot shows recorded positions.")
```

Orange lines are my movements. A rock festival to the east of Antwerp, some visits to the coast, to my home city Ghent, a marriage in the west, biking and text mining in Leuven, and hikes in the woods of the south. Looks about right. And yes, I live and work in Brussels.

So let's take a closer look at Brussels.

```{r message=FALSE}
brussels <- get_map(location = 'Brussels', zoom = 12, maptype = "terrain-lines")

```

```{r fig.height=10, fig.width=15}
bb <- attr(brussels, "bb")
bb
```
```{r}
# just keeping data from brussels

x <- df %>% 
  filter(lat < bb$ur.lat & lat > bb$ll.lat,
         lon > bb$ll.lon & lon < bb$ur.lon)
```



```{r fig.height=10, fig.width=10}
ggmap(brussels) + 
  geom_point(data = x, aes(x = lon, y = lat),
             alpha = 0.5, color = "#FC4E07", size = .65) + 
  theme(legend.position = "right") + 
  labs(
    x = "Longitude", 
    y = "Latitude", 
    title = "Location history data points in Brussels",
    caption = "\nA simple point plot shows recorded positions.")

```

Let's see how much we can zoom in.

```{r message=FALSE}

zoom <- get_map(location = c(lon = 4.349078, lat = 50.850586), zoom = 16, maptype = "satellite")


```

```{r}
bb <- attr(zoom, "bb")
x <- df %>% 
  filter(lat < bb$ur.lat & lat > bb$ll.lat,
         lon > bb$ll.lon & lon < bb$ur.lon)
```


```{r fig.height=9, fig.width=9}

ggmap(zoom) + 
 geom_point(data = x, aes(x = lon, y = lat),
             alpha = 0.7, color = "#FC4E07", size = .85) + 
  theme(legend.position = "right") + 
  labs(
    x = "Longitude", 
    y = "Latitude", 
    title = "Location history data points in central Brussels",
    caption = "\n favorite haunts.")
    
```

So that is pretty consistent with my favorite downtown haunts. Let's pan out and add the number of seconds from 6 am in the day to the dataframe. 

```{r}
period_to_seconds(hms("06:00:00")) # seconds at 6am

period_to_seconds(hms("23:59:59")) # seconds just before midnight
```

```{r}

df <- df %>% 
  mutate(from_dawn = 
           case_when((period_to_seconds(hms(time)) > 21600) ~ 
                       (period_to_seconds(hms(time))-21600),
                     TRUE ~ (period_to_seconds(hms(time))+64799)))


```


So now there is a value (from_dawn) that is equal to 0 at six am and continues to increase until just before six. Easy to convert to hours starting from six am. Values seem higher in the center of the city.  "It's six o'clock in the morning, do you know where your parents are?" :waxing_gibbous_moon:  

```{r fig.height=10, fig.width=10, warning=FALSE}
options(stringsAsFactors = T)

bb <- attr(brussels, "bb")
x <- df %>% 
  filter(lat < bb$ur.lat & lat > bb$ll.lat,
         lon > bb$ll.lon & lon < bb$ur.lon) %>% 
  mutate(hours = from_dawn/3600)

ggmap(brussels) + 
  stat_summary_2d(geom = "tile", bins = 100, data = x, 
                  aes(x = lon, y = lat, z = hours), alpha = 0.7) + 
  scale_fill_gradientn(colors  =  viridis(4), 
                       guide = guide_legend(title = "hours since 6am")) +
  labs(
    x = "Longitude", 
    y = "Latitude", 
    title = "Location history data points around Brussels",
    subtitle = "Color scale shows accuracy (low: blue, high: red)",
    caption = "\nThis bin plot shows recorded positions 
    and their time of day")
```


Looking at it by year.


```{r fig.height=10, fig.width=10}

bb <- attr(brussels, "bb")
x <- df %>%
  filter(lat < bb$ur.lat & lat > bb$ll.lat,
         lon > bb$ll.lon & lon < bb$ur.lon,
         year > 2014)

ggmap(brussels) + facet_wrap(~year) +
geom_point(data = x, aes(x = lon, y = lat), alpha = 0.5, color = "#FC4E07", size = .8) + 
  theme(legend.position = "right") + 
  labs(
    x = "Longitude", 
    y = "Latitude", 
    title = "Location history data points in Brussels - by year ",
    caption = "\nA simple point plot shows recorded positions.")
```

It would seem that I moved a lot in 2016 compared to the other years, but to that extent it does not look right. Let's dig a little deeper in the data.

```{r fig.height=10, fig.width=10}

ggplot(data = x, aes(x = datetime, y = lat))+
  geom_line()+
  theme(axis.title.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

```
So clearly 2015 and 2017 have much less data. In 2017 it looks like I only used location services to three or four specific places in Brussels, but it is intriguing that there seems to be data throughout the year. And this in contrast with the end of 2018 when I switched the location tracking off. 

Whether Google still keeps track of my whereabouts is of course another matter. 

I still don't know if my commute has gotten longer. I will try to find out in a next blog post.

```{r}

```

