---
title: members of the chamber of representatives of Belgium, 55th legistlature
author: William Bourgeois
date: '2019-07-16'
slug: members-of-the-chamber-of-representatives-of-belgium-55th-legistlature
categories: []
tags:
  - scraping
  - Belgian politics
  - Twitter
---

On 20/06/2019 the new chamber of representative was constituded for the 55th legistlature. There are a fair number of new members in this assembly. In this post we will get their Twitter handles if they have one. This will make it possible to get their tweets and information about their accounts to be used in various anlyses. 

The process can get a tiny bit tidious because as far as I know there is no website regrouping the handles, we need to search for them using the rtweet package, roughly in the same way in which it was done for the last legistlature. 

The names of the new parliamentarians have been published on the official website (dekamer.be), but a lot of pictures are still missing so let's focuss on Twitter and get a list of names first.


```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("rvest")
library("scales")
library("rtweet")
#library("glue")

```



```{r}

url <- paste0("https://www.dekamer.be/kvvcr/showpage.cfm?section=/",
              "depute&language=nl&cfm=cvlist54.cfm?legis=55&today=y") # url to the current members

table <- url %>% 
  read_html() %>% 
  html_nodes("#story > table") %>% 
  html_table()

table <- table[[1]] # extracting the dataframe

names(table) <- c("ln_fn", "party", "d1", "d2")

reps_55<- table %>% 
  select(ln_fn, party) %>% 
  arrange(ln_fn)

head(reps_55) # have a peek

```


Some members were also part of the last legistlature, and their handles were already found in a past post.


```{r}
twitter_users <- readRDS("~/R/blogs/content/post/data/20190502/twitter_users.rds") # the pre-exiting data

handles <- twitter_users %>% 
  select(ln_fn, screen_name)

missing_handles <- reps_55 %>% 
  left_join(handles) %>% 
  filter(is.na(screen_name))

```

```{r}
nrow(missing_handles)
```

So of the 150 current members, we are missing 83 handles. Makes you wonder how much current members did not seat in the last legistlature...

```{r message=FALSE, warning=FALSE}
url <- paste0("https://www.dekamer.be/kvvcr/showpage.cfm?section=/depute&language=nl&cfm=cvlist54.cfm?legis=54&today=n") # url to the members of 54

table <- url %>% 
  read_html() %>% 
  html_nodes("#story > table") %>% 
  html_table()

table <- table[[1]] # extracting the dataframe

names(table) <- c("ln_fn", "party", "d1", "d2")

reps_54<- table %>% 
  select(ln_fn) %>% 
  arrange(ln_fn) %>% 
  mutate(legist = 54)

newer_members <- reps_55 %>% 
  left_join(reps_54) 


head(newer_members) # have a peek

```

Of course joining on names is not ideal, since this is error prone, but the results look more or less correct and the method will give a ball park figure to work with.

```{r}
newer_members %>% filter(is.na(legist)) %>% 
  nrow()/150
```

So that is 55% members who did not seat in the past legistlature. What are the parties with the most new members? 

```{r}

newer_members %>% filter(is.na(legist)) %>% 
  group_by(party) %>% 
  summarise(new = n()) %>% 
  arrange(desc(new))

```

And what are the parties with the highest ratio of new members?

```{r message=FALSE, warning=FALSE}
newer_members %>%  
  group_by(party) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total)) %>% left_join(newer_members
                                     %>% filter(is.na(legist)) %>% 
                                       group_by(party) %>% 
                                       summarise(new = n())) %>% 
  mutate(prct = percent(new/total)) %>% # using scales package
  arrange(desc(prct))

```

# check because the 100% are here on the bottom?

Interesting. Of course the VB and PVDA-PTB have very high ratios, but so has the sp.a and others. But the point was to get the Twitter handles. 

A function can be created to tweek the search_user function of the rtweet pachage. This tweak also limits the result to one, so that most obvious ones can be harvested. 

```{r}
search_one <- function(x){print (x)
  search_users(x, n=1) %>% # only return one line per name
    mutate(ln_fn = x)} # keep track of who we searched for
```



```{r message=FALSE}
twitter_users_new <- map_df(missing_handles$ln_fn,possibly(.f= search_one, otherwise = NULL))
```

The result contains 89 variables, so let's looks at only some of them to see if we catched some reps Twitter handles. 

```{r}
twitter_users_new %>% 
  select(name, ln_fn, description, text, screen_name)

```

