---
title: r u bot or not bot?
author: William Bourgeois
date: '2019-08-20'
slug: r-u-bot-or-not-bot
categories: []
tags:
  - Twitter
---

Yesterday Twitter announced that they had identified 936 accounts originating from China that were instrumentalised for 'information operations' in Hong Kong. These were 'deliberately and specifically attempting to sow political discord in Hong Kong, including undermining the legitimacy and political positions of the protest movement on the ground'. 

I find it interesting to know why Twitter decided to tackle 'information operations' targeting protest actions in Hong Kong. 

The couple of screenshots and the four files that Twitter let's us see, do not illustrate in any way that these accounts were involved in 'coordinated inauthentic behavior', from a 'state-backed' operation.  

I'm not saying it was a bad decision to suspend the accounts, what I'm saying is that Twitter's criteria are not transparent and that the information they provide do not prouve that a state actor was involved. 

They state to have identified 'large clusters of accounts behaving in a coordinated manner to amplify messages related to the Hong Hong protests.' Well I hope that the Hong Kong protesters are behaving in exactly the same way. 

Who decides when accounts are behaving in a legitimate matter or not? Using which hard criteria? Hard ones and not some woolly [ones](https://help.twitter.com/en/rules-and-policies/platform-manipulation) By these criteria would the account of the orange menace not need to be suspended? Does that account not 'attempting to sow political discord'? Who decides? How?

Twitter states:'These deceptive strategies have been around for far longer than Twitter has existed.' Deceptive strategies, e.g. propaganda have been around indeed for longer than Twitter existed. And so has censure. 

In this, the largest of its kind in the industry, 'information operation', much more transparency from Twitter is required. 

You can find the full statement [here](https://blog.twitter.com/en_us/topics/company/2019/information_operations_directed_at_Hong_Kong.html)


Here are the accounts:

```{r}
library(readr)
china_2 <- read_csv("./data/20190820/china_082019_2_users_csv_hashed.csv")
china_1 <- read_csv("./data/20190820/china_082019_1_users_csv_hashed.csv")
china <- rbind(china_1, china_2)
```

And their columns:

```{r}
names(china)
```

The tweets:

```{r}
china_2_tweets <- read_csv("./data/20190820/china_082019_2_tweets_csv_hashed.csv")
china_1_tweets <- read_csv("./data/20190820/china_082019_1_tweets_csv_hashed.csv")
china_tweets <- rbind(china_1_tweets, china_2_tweets)
```

The columns of the tweets:
```{r}

names(china_tweets)
```


```{r}
library("tidyverse")
user_profile_description <- china_tweets %>% 
  select(user_profile_description) %>% 
  unique()

```


What is the point in sharing data if it does not prove anything when you scratch the surface? Is this transparancy washing?

Chua Chin Hon goes much further in the analasys of the propaganda campaign and I recommend reading his excellent series of [posts](https://towardsdatascience.com/failed-surge-analyzing-beijings-disinformation-campaign-surge-on-twitter-b376aed812ed) on the issue.

He also points to data that I missed or was published later. AO 32 zipped files like this one: https://storage.cloud.google.com/twitter-election-integrity/hashed/2019_08/china_082019_1/china_082019_1_tweet_media_hashed/china_082019_1_tweet_media_hashed_001.zip

But these seem just to be a dump of the media sent with the tweets (?). The first one contains a lot of videos of US sports clips. Without meta-data it makes even less sence. I don't understand why this data was posted. Data without context is just noise.  

<video width="480" height="320" controls="controls">
  <source src="./data/20190831/sample.mp4" type="video/mp4">
</video>

-> there is a package that aims to indetify bots
-> two major bugs needed to be fdixed
-> learning was cool
-> results are shit anyway
-> all the more need for Twitter transparancy

But maybe I am the one that is missing something. So I'll end the rant here.



In this instance the accounts are said to be trolls. 