---
title: Belgian parliamentarians Twitter accounts
author: William Bourgeois
date: '2019-05-02'
slug: belgian-parliamentarians-twitter-accounts
categories: []
tags:
  - Belgian politics
  - Twitter
---

Now that we can scrape the names of the parliamentarians of the 54th legislature, we can try to find their Twitter accounts too. 

Let's start by reconstructing the list. Then we can search Twitter for their accounts handles.  A bit of grunt work and no visuals in this post, but it is needed to harvest parlaimentarians' tweets soonish. 

Starting with attaching libraries, and in particular ['rtweet'](https://rtweet.info/). 


```{r echo=TRUE, message=FALSE, warning=FALSE}
#libraries
library("tidyverse")
library("rvest")
library("glue")
library("rtweet")
library("purrr")

# the ones seating currently, with their (current) party

url <- paste0("http://www.dekamer.be/kvvcr/showpage.cfm?section=",
              "/depute&language=nl&cfm=/site/wwwcfm/depute/cvlist54.cfm")

table <- url %>% 
  read_html() %>% 
  html_nodes("#story > table") %>% 
  html_table()

table <- table[[1]]

names(table) <- c("ln_fn", "party", "d1", "d2") # ln_fn is last name, first name
table <- table %>% 
  select(ln_fn, party)

table <- table %>%
  mutate(ln_fn = str_replace_all(ln_fn,"  +"," "))

# the ones that have left (without party)

url <- paste0("https://www.dekamer.be/kvvcr/showpage.cfm?section",
              "=/depute&language=nl&cfm=cvlist54.cfm?legis=54&today=n")

table_54 <- url %>% 
  read_html() %>% 
  html_nodes("#story > table") %>% 
  html_table()

table_54 <- table_54[[1]] # extracting the dataframe

names(table_54) <- c("ln_fn", "d1", "d2", "d3")  
table_54 <- table_54 %>% 
  select(ln_fn) %>% 
  arrange(ln_fn)

table_54 <- table_54 %>%
  mutate(ln_fn = str_replace_all(ln_fn,"  +"," "))

old_members <- table_54 %>% 
  anti_join(table) %>% 
  mutate(party = "tbd")

table <- rbind(table, old_members)

saveRDS(table, "./data/20190502/table.rds")

rm(old_members)
rm(table_54)
rm(url)
```
How many do we have?

```{r}
nrow(table)
```

Ok, did not lose any along the way. Now we have a data frame with the last name and first name of the 179 persons who seated in the 54th legislature (by now there will be few or no changes, since we are so close to this month's elections). 

Next the twitter API credentials are needed to use the rtweet package. Documentation can be found [here](https://rtweet.info/#api-authorization)

Once authenticated, based on our parliamentarian's name and using the 'search_users' function we can find their twitter accounts. I had some success using a for loop in the past, but will try to use the purrr package this time around.

```{r}
# so tweak the function, because it tries to find by default 100 users
# and because we want to keep track of who we searched for in the results
# take only first result for now

search <- function(x){print (x)
  search_users(x, n=1) %>% # only return one line per name
    mutate(ln_fn = x)} # keep track of who we searched for

```

Using the function 'possibly' a data frame can be created without stopping or including errors.

```{r echo=TRUE, message=FALSE, warning=FALSE, results=FALSE}
twitter_users <- map_df(table$ln_fn,possibly(.f= search, otherwise = NULL))
```

The nice thing about it is that a lot of information can be collected using the Twitter API. Have look yourself: 
```{r}
names(twitter_users)
```

Some of the fields, like "hashstags" contain lists.

Arranging columns:

```{r}
twitter_users <- twitter_users %>% 
  select(ln_fn, everything())
```

These names return an error using the "twitter_user" function. But these politicians do not tweet. We can skip them.

```{r}
list_no_twitter <- c("Detiège Maya", "Mathot Alain", "Pivin Philippe",
                     "Wilrycx Frank")
```

We now have a data frame with 152 users for our 175 politicians. 

We need to find the right user ids or twitter handles, by looking at key fields in the data frame returned by the 'search_user' function. (summarised in a smaller data frame)

```{r}
inspect <- twitter_users %>%
  select(ln_fn, name, screen_name, description, text)
```

Luckily most often the first result is the correct one. Although there are also sometimes surprises. At least these do not seem like descriptions of Belgian parliamentarians.

```{r}

descriptions <- inspect %>% 
  filter(ln_fn == "Bracke Siegfried" 
         | ln_fn == "Fernandez Fernandez Julie"
         | ln_fn == "Top Alain"
         | ln_fn == "Bonte Hans") %>% 
  select(description)

descriptions

```

These seem obviously wrong. 

```{r}

find_again <- tibble("ln_fn" = c("Bonte Hans","Ceysens Patricia","Chabot Jacques",
                                 "Fernandez Fernandez Julie","Geerts David",
                                 "Henry Olivier","Janssen Werner", "Janssens Dirk",
                                 "Matz Vanessa", "Metsu Koen", "Miller Richard",
                                 "Pehlivan Fatma","Smeyers Sarah","Top Alain",
                                 "Van de Velde Robert"))
```

Would they be first 10 of the search result?
```{r message=FALSE}
search <- function(x){print (x)
  search_users(x, n=10) %>% 
    mutate(ln_fn = x)} # keep track of who we searched for

twitter_users_2 <- map_df(find_again$ln_fn,
                          possibly(.f= search, otherwise = NULL))

inspect <- twitter_users_2 %>%
  select(ln_fn, name, screen_name, description, text)

```

The new batch can easily be identified and added to the first batch.

```{r}
twitter_users <- twitter_users %>% 
  filter(!ln_fn %in% find_again$ln_fn) %>% 
  rbind(twitter_users_2 %>% 
  slice(13, 21, 30, 45, 51, 71, 76 , 99, 113))
```

We are still missing these:

```{r}
setdiff(setdiff(table$ln_fn, twitter_users$ln_fn), list_no_twitter)
```

Some of these surely have a twitter account. More searching is needed to find: 

@CarolineCassart
@patriciaceysens
@podelannois
@leendierick
@dirkjanssens19
@millerrichardmr
@nathaliemuylle
@jefvandenbergh
@VincentVQ
@VvelthovenPeter
@pehlivan_fatma
@FrankWilrycx
@McMarghem
@robvandevelde

```{r}
more <- tibble(handle = wrapr::qc(CarolineCassart, patriciaceysens, podelannois,
                                  leendierick,dirkjanssens19, millerrichardmr,
                                  nathaliemuylle, jefvandenbergh, VincentVQ,
                                  VvelthovenPeter, pehlivan_fatma, FrankWilrycx,
                                  McMarghem),
               ln_fn = c("Cassart-Mailleux Caroline", "Ceysens Patricia",
                         "Delannois Paul-Olivier", "Dierick Leen", 
                         "Janssens Dirk", "Miller Richard", "Muylle Nathalie", 
                         "Van den Bergh Jef", "Van Quickenborne Vincent", 
                         "Vanvelthoven Peter", "Pehlivan Fatma" , 
                         "Wilrycx Frank", "Marghem Marie-Christine"))

more_1 <- more %>% 
  filter(handle !="pehlivan_fatma",
         handle !="FrankWilrycx")

more_2 <- more %>% 
  filter(handle =="pehlivan_fatma" |
         handle =="FrankWilrycx")

```



```{r message=FALSE}
search <- function(x){print (x)
  search_users(x, n=1)} # keep track of who we searched for

twitter_users_3 <- map_df(more_1$handle,possibly(.f= search, otherwise = NULL)) %>% 
  cbind(more_1$ln_fn) %>%
  rename("ln_fn" = "more_1$ln_fn") %>% 
  select(ln_fn, everything())

twitter_users <- rbind(twitter_users, twitter_users_3)

nrow(twitter_users)

```

Still problems with the last two. "FrankWilrycx" probably because he has not tweeted yet. And "pehlivan_fatma" was a bit harder to find


```{r echo=TRUE, message=FALSE}

more_2 <- more %>% 
  filter(handle =="pehlivan_fatma")

search <- function(x){print (x)
  search_users(x, n=11)} # keep track of who we searched for

twitter_users_4 <- map_df(more_2$handle,
                          possibly(.f= search, 
                                   otherwise = NULL)) %>% 
  cbind(more_2$ln_fn) %>%
  rename("ln_fn" = "more_2$ln_fn") %>% 
  select(ln_fn, everything()) %>% 
  slice(11)

twitter_users <- rbind(twitter_users, twitter_users_4)

nrow(twitter_users)

```


There are 159 Twitter handles, out of 179 parliamentarians. 

(if you want to have fun with the profile pics they are in https://twitter.com/{handle}/photo)

We are now ready to extract a considerable amount of their tweets. 

```{r}

saveRDS(twitter_users, "./data/20190502/twitter_users.rds")

```


