---
title: Topic modeling Facebook ads
author: William Bourgeois
date: '2020-03-03'
slug: topic-modeling-facebook-ads
categories: []
tags:
  - Facebook
  - nlp
  - Belgian politics
---

In the last two posts there was an analysis of Facebook ads of Belgian politicians or political parties. To temporary conclude the subject we will do some topic modeling on these ads.

```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("cld2")
library("tidytext")
library("topicmodels")
library("qdap")
```

Loading the data.

```{r}
df <- readRDS("./data/20200104/df.rds")
# create a doc_id
df <- rowid_to_column(df, "doc_id")
```


I think that one of the first things to do is to is to determine the languages the ads were written in. But since a number of the ads are writen in more than one language it might be better to first parse them by sentences, while identifying the sentences still that belong to the same ad. 

```{r eval=FALSE, include=TRUE}
sentences <- data.frame()
for (i in (1:nrow(df))){
  sentence_extract <- t(str_extract_all(df$ad_creative_body[i], boundary("sentence"), simplify = TRUE)) %>% 
    as.data.frame()
  sentence_extract <- bind_cols("sentences" = sentence_extract, "doc_id" = rep.int(df$doc_id[i],
                                                                                   nrow(sentence_extract)))
  sentences <- rbind(sentences, sentence_extract)
}

```

```{r, include=FALSE}
#saveRDS(sentences, "./data/20200303/sentences.rds")
sentences <- readRDS("./data/20200303/sentences.rds")
```

Did it work?

```{r}
sentences
```

Indeed.

So now we can filter and take only the dutch sentences. And then we can tokenize to words (while transforming CD&V, sp.a and N_VA in order to keep them as tokens).

Download stopwords
```{r}
stop <- read.delim("http://www.damienvanholten.com/downloads/dutch-stop-words.txt", header=FALSE)

sentences_cld2 <- sentences %>%
  mutate(clean_text = as.character(V1)) %>% 
  mutate(language = detect_language(clean_text)) 

sentences_nl <- sentences_cld2 %>% 
  filter(language == "nl") %>% 
  select(-V1, -language) %>% 
  distinct(clean_text, .keep_all = TRUE)
# remove stopwords

tm::removeWords("en ik alles behalve maar in", tm::stopwords("dutch"))
tm::removeWords("and", tm::stopwords("english"))


sentences_nl <- sentences_nl %>% 
  mutate(clean_text = tolower(clean_text)) %>% 
  mutate(clean_text = tm::removeWords(clean_text, c(tm::stopwords("dutch"), 
                                                    "én", "http", "https","www.s", 
                                                    "bit.ly", "wwww", "pagina")))

sentences_nl <- sentences_nl %>% 
  mutate(clean_text = str_replace_all(clean_text, "sp.a", "spxxxxa"),
         clean_text = str_replace_all(clean_text, "CD&V", "CDxxxxV"),
         clean_text = str_replace_all(clean_text, "Cd&v", "CDxxxxV"),
         clean_text = str_replace_all(clean_text, "cd&v", "CDxxxxV"),
         clean_text = str_replace_all(clean_text, "vlaams belang", "VlaamsxxxxBelang"),
         clean_text = str_replace_all(clean_text, "open vld", "Openxxxxvld"),
         clean_text = str_replace_all(clean_text, "n-va", "NVxxxxA"))

# sp.a , cd&v and n-va 

by_doc_word <- sentences_nl %>%
  select(doc_id, clean_text) %>% 
  unnest_tokens(words, clean_text, token = "words")

by_doc_word_2 <- sentences_nl %>%
  select(doc_id, clean_text) %>% 
  unnest_tokens(words, clean_text, token = "ngrams", n = 2)

by_doc_word_3 <- sentences_nl %>%
  select(doc_id, clean_text) %>% 
  unnest_tokens(words, clean_text, token = "ngrams", n = 3)


by_doc_word <- by_doc_word %>% 
  rbind(by_doc_word_2, by_doc_word_3)
#unnest_tokens(words, clean_text)

```


```{r}
names(stop) <- "words"
word_counts <- by_doc_word %>% 
  anti_join(stop) %>% 
  count(doc_id, words, sort = TRUE) %>% 
  filter(nchar(words) < 27) %>% 
  filter(nchar(words) > 1)
word_counts
```

Let's see if we have corectly tokenised in words:
```{r}

head(by_doc_word %>% 
       filter(doc_id > 1000),10)
```

And now we can make a document term matrix. With as its name suggests is a matrix with one row per ad (document) and one column per word here populated by the frequency of the word in the ad.


```{r}
doc_dtm <- word_counts %>%
  cast_dtm(doc_id, words, n, weighting = tm::weightTf)
doc_dtm
```


Let's start finding some topics.

```{r}
# 
n_topics <- 25
doc_lda <- LDA(doc_dtm, k = n_topics, method = "Gibbs", 
               control = list(iter=2000, seed=42,alpha = 0.4))
doc_terms <- as.data.frame(topicmodels::terms(doc_lda, 15), stringsAsFactors = FALSE)
doc_terms[1:n_topics]

```

```{r}
doc_topics <- topicmodels::topics(doc_lda, 1)
doc_topics_df <- as.data.frame(doc_topics)

theta <- as.data.frame(topicmodels::posterior(doc_lda)$topics)
head(theta[1:5])
```


Trying library(textmineR) for topic modeling. Here is some background [info](https://cran.r-project.org/web/packages/textmineR/vignettes/c_topic_modeling.html)

```{r}
library("textmineR") #install.packages("textmineR")
```

```{r}
documents <- sentences_nl %>%
  group_by(doc_id) %>%
  mutate(text = paste0(clean_text, collapse = "")) %>%
  select(doc_id, text) %>% 
  unique()

# only take texts from biggest flemish parties
top_seven <- documents %>% 
  left_join(df) %>% 
  group_by(name_party) %>%
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  top_n(7) %>% 
  select(name_party)

documents <- documents %>% 
  left_join(df) %>% 
  filter(name_party %in% top_seven$name_party) %>% 
  select(doc_id, text)
  
  
  
```


```{r}
dtm <- CreateDtm(doc_vec = documents$text, # character vector of documents
                 doc_names = documents$doc_id, # document names
                 ngram_window = c(1, 2), # minimum and maximum n-gram length
                 stopword_vec = c(stopwords::stopwords("nl"), # stopwords from tm
                                  stopwords::stopwords(source = "smart")), # this is the default value
                 lower = TRUE, # lowercase - this is the default value
                 remove_punctuation = TRUE, # punctuation - this is the default
                 remove_numbers = TRUE, # numbers - this is the default
                 verbose = FALSE, # Turn off status bar for this demo
                 cpus = 2) # default is all available cpus on the system

dtm <- dtm[,colSums(dtm) > 2]
```


```{r}
set.seed(42)
n_topics <- 30

model <- FitLdaModel(dtm = dtm, 
                     k = n_topics,
                     iterations = 200, 
                     burnin = 175,
                     alpha = 0.08,
                     beta = 0.05,
                     optimize_alpha = TRUE,
                     calc_likelihood = TRUE,
                     calc_coherence = TRUE,
                     calc_r2 = TRUE,
                     cpus = 2) 
```

```{r}
str(model)
```
```{r}
# R-squared 
# - only works for probabilistic models like LDA and CTM
model$r2
#> [1] 0.2747765

```

```{r}
# log Likelihood (does not consider the prior) 
ggplot(data = model$log_likelihood, aes(x= iteration, y= log_likelihood))+
  geom_line()+
  labs(title="log likelihood",
        x ="iterations", y = "log likelihood")+
  theme_minimal()
```


We can analyse the models probabilistic coherence measures of topics that are defined as "Topic Coherence measures score a single topic by measuring the degree of semantic similarity between high scoring words in the topic. These measurements help distinguish between topics that are semantically interpretable topics and topics that are artifacts of statistical inference..." by the authors in the paper [here](https://www.aclweb.org/anthology/D12-1087.pdf) 


```{r}
# probabilistic coherence, a measure of topic quality
# this measure can be used with any topic model, not just probabilistic ones
model$coherence
summary(model$coherence)
#>    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#>  0.0060  0.1188  0.1543  0.1787  0.2187  0.4117
```

https://datascienceplus.com/evaluation-of-topic-modeling-topic-coherence/


```{r}
hist(model$coherence, 
     col= "blue", 
     main = "Histogram of probabilistic coherence")
```

```{r}
# hist(model$coherence, 
#      col= "blue", 
#      main = "Histogram of probabilistic coherence")

df <- as.data.frame(model$coherence) %>% 
  rownames_to_column("topic") %>% 
  mutate(topic = str_remove(topic, "t_"),
         topic = as.numeric(topic)) 
names(df) <- c("topic", "coherence")

ggplot(data = df, aes(x=topic,y=coherence, fill=topic))+
  geom_bar(stat = "identity",  color="blue")+
  theme_minimal()
```


So this is what the probabilistic coherence of our topics look like using our parameters and 30 topics. 

We could cylce through a range of number of topics to see what number of topics gives the best coherence score.

```{r}
results <- data.frame()
for (i in (5:45)){
  print(i)
  n_topics <- i

  model <- FitLdaModel(dtm = dtm, 
                     k = n_topics,
                     iterations = 200, 
                     burnin = 175,
                     alpha = 0.08,
                     beta = 0.05,
                     optimize_alpha = TRUE,
                     calc_likelihood = TRUE,
                     calc_coherence = TRUE,
                     calc_r2 = TRUE,
                     cpus = 2) 
  
  r2 <- model$r2
  row <- data.frame("number of topics" = i, "r2" = r2)
  results <- rbind(results, row)
  
}

```




```{r}
results2 <- data.frame()
for (i in (25:45)){
  print(i)
  n_topics <- i

  model <- FitLdaModel(dtm = dtm, 
                     k = n_topics,
                     iterations = 200, 
                     burnin = 175,
                     alpha = 0.08,
                     beta = 0.05,
                     optimize_alpha = TRUE,
                     calc_likelihood = TRUE,
                     calc_coherence = TRUE,
                     calc_r2 = TRUE,
                     cpus = 2) 
  
  r2 <- model$r2
  print(r2)
  row <- data.frame("number of topics" = i, "r2" = r2)
  results2 <- rbind(results2, row)
  
}

```

```{r}
ggplot(data = results2, aes(x = , y=))
```


```{r}
# Get the top terms of each topic
(model$top_terms <- GetTopTerms(phi = model$phi, M = 10))
```

```{r}
# Get the prevalence of each topic
# You can make this discrete by applying a threshold, say 0.05, for
# topics in/out of documents. 
model$prevalence <- colSums(model$theta) / sum(model$theta) * 100

# prevalence should be proportional to alpha
plot(model$prevalence, model$alpha, xlab = "prevalence", ylab = "alpha")
```

```{r}
# textmineR has a naive topic labeling tool based on probable bigrams
model$labels <- LabelTopics(assignments = model$theta > 0.05, 
                            dtm = dtm,
                            M = 1)

head(model$labels, 10)
```

```{r}
# put them together, with coherence into a summary table
model$summary <- data.frame(topic = rownames(model$phi),
                            label = model$labels,
                            coherence = round(model$coherence, 3),
                            prevalence = round(model$prevalence,3),
                            top_terms = apply(model$top_terms, 2, function(x){
                              paste(x, collapse = ", ")
                            }),
                            stringsAsFactors = FALSE)
```

```{r}
model$summary[ order(model$summary$prevalence, decreasing = TRUE) , ][ 1:30 , ]
```



```{r}

theta <- model$theta %>% 
  as.data.frame()

max <- apply(theta[, 1:n_topics], 1, max) %>% 
  as.data.frame() %>% 
  rename("max" = ".")

max$doc_id = row.names(max)
  
```

```{r}

theta$topic <- colnames(theta)[apply(theta,1,which.max)]
theta$doc_id = row.names(theta)
```

```{r}
class(documents$doc_id)
```


```{r}
test_ana <- theta %>% 
  mutate(doc_id = as.integer(doc_id)) %>% 
  left_join(max %>% 
              mutate(doc_id = as.integer(doc_id))) %>% 
  left_join(documents) %>% 
  left_join(df) %>% 
  select(doc_id, max, topic, text, name_party) %>% 
  filter(max > 0.599)
```

```{r}
y <- test_ana %>% group_by(name_party, topic) %>% 
  summarise(sum_n = n())

highest_seven <- y %>% group_by(name_party) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count)) %>% 
  top_n(7)
highest_seven
z <- y %>% 
  mutate(topic = str_remove_all(topic, "t_")) %>% 
  mutate(topic = as.integer(topic)) %>% 
  filter(name_party %in% highest_seven$name_party) %>% 
  ungroup()
```

```{r}
  
ggplot(data = z, aes(x = topic, y = sum_n)) +
  geom_bar(stat = 'identity', width = 0.65)+
  facet_wrap(~name_party, scales = 'free')+
  theme(axis.text.x = element_text(angle = 90))+
  theme_minimal()+
  ylab("number of ads")+
  aes(fill = as.factor(name_party))+
  theme(legend.position = "none")+
  scale_fill_manual(values = c("n-va"= "sienna3", 
                                        "vlaams belang" = "grey35", 
                                        "groen" = "lightgreen",
                                        "open vld" = "royalblue2",
                                        "cdh" = "orange2", 
                                        "ps" = "pink", 
                                        "sp.a" = "pink",
                                        "pvda" ="firebrick2",
                                        "ptb" = "firebrick2",
                                        "cd&v" = "orange"
                                        ))
```


```{r}
z %>% 
  filter(name_party == "vlaams belang") %>% 
  filter(topic == 28)

```

```{r}
test_ana %>% 
  filter(name_party == "vlaams belang") %>%
  filter(topic == "t_28") %>% 
  arrange(desc(max)) %>% 
  head(8) %>% 
  select(text)
  
```

```{r}
z %>% 
  filter(name_party == "pvda") %>% 
  filter(topic == 27)
```


```{r}
test_ana %>% 
  filter(name_party == "pvda") %>%
  filter(topic == "t_27") %>% 
  arrange(desc(max)) %>% 
  head(8) %>% 
  select(text)
```

```{r}
z %>% 
  filter(name_party == "n-va") %>% 
  filter(topic == 3)

```
```{r}
test_ana %>% 
  filter(name_party == "n-va") %>%
  filter(topic == "t_3") %>% 
  arrange(desc(max)) %>% 
  head(8) %>% 
  select(text)
```


